From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-07-22 dictionary.tex

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \labitem{(1)}{c1} $\TrFrob{G} : \CSiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \labitem{(2)}{c2} every character sheaf on $G$ is isomorphic to one defined by a discrete isogeny;
 \labitem{(3)}{c3} $\Aut(\cs{L}) = 1$, for all character sheaves $\cs{L}$ on $G$.
 \end{enumerate}
 \end{proposition}
\begin{proof}
By Lemma~\ref{lem:bounded_connected}, we know that every character sheaf $\cs{L}$ on $\bG$ descends to $G$;
let $\cs{E}$ be an object of $\bCS(G)$ for which $B_G(\cs{E}) \iso \cs{L}$.
 Since the functor $B_G : \bCS(G) \to \CS(G)$ is full and faithful, $\Aut(\cs{L}) = \Aut(\cs{E})$.
From here, Deligne's function-sheaf dictionary for connected commutative algebraic groups over finite fields,
as in \cite{deligne:SGA4.5}*{Sommes trig.} or \cite{laumon:87a}*{1.1.3}, gives us all we need for points \ref{c1} and \ref{c2}.
% as we briefly recall.
%\todo{Fix citelist} %\citelist{\cite{deligne:SGA4.5}*{Sommes trig.},\cite{laumon:87a}*{1.1.3}}
\iffalse
As in the proof of Proposition~\ref{prop:finite}, use \'etale descent to see that pullback by the Lang isogeny $\Lang : G\to G$
defines an equivalence of categories between local systems on $G$ and $G(\Fq)$-equivariant local systems on $G$. 
Under this equivalence, local systems $\cs{E}$ on $G$ arising from objects in $\bCS(G)$ are matched with $G(\Fq)$-equivariant
constant local systems of rank-one on $G$, and therefore with one-dimensional representations of $G(\Fq)$. 
In the same way, pullback along the isogeny $\Lang\times\Lang : G\times G\to G\times G$ matches the extra structure
$\mu_\cs{E} : m^*\cs{E} \to \cs{E}\boxtimes\cs{E}$ with an isomorphism $m^*V \to V\boxtimes V$ of one-dimensional
representations of $G(\Fq)\times G(\Fq)$, which is exactly an isomorphism $V \to V\otimes V$ of one-dimensional representations,
which is exactly the choice of an isomorphism $V\iso \EE$.
We see that $\bCS(G)$ is equivalent to the category of characters of $G(\Fq)$.
Let $\bg$ be a geometric point above $g \in G(\Fq)$.  If $\cs{E}$ matches $\psi : G(\Fq)\to \EEx$ under this equivalence,
a simple calculation on stalks reveals that the action of Frobenius on $\cs{E}_{\bg}$ is multiplication by $\psi(g)^{-1}$.
In other words, for every $\cs{E}$ in $\bCS(G)$, the trace of $\Lang^*\cs{E}$ is $\trFrob{\cs{E}}^{-1}$ as a
representation of $G(\Fq)$, proving parts \ref{c1} and \ref{c2}.
\fi
For part \ref{c3}, suppose $\Lang^*\cs{E} = V$ with isomorphism $V \to V\otimes V$.
Observe that the equivalence above establishes a bijection between $\Aut(\cs{E})$ and the group of automorphisms of $\rho : V\to V$ for which 
\[
\begin{tikzcd}
\arrow{d}{} V \arrow{r}{\rho} & V\arrow{d}{}\\
V\otimes V \arrow{r}{\rho\otimes \rho} & V\otimes V
\end{tikzcd}
\]
commutes. 
Since the only such isomorphism $\rho$ is $\id_V$, it follows that $\Aut(\cs{E}) = 1$, completing the proof.
\end{proof}

@@

%\begin{remark}
%Although we will mainly use the group homomorphism $S_G : \CSiso{G}\to \Hh^2(E^\bullet_G)$ when $G$
%is an \'etale commutative group scheme, the function is defined for all smooth commutative group schemes $G$;
%however, the next result does not extend to all smooth commutative group schemes.
%\end{remark}

@@

In this section we fit the group homomorphisms $\TrFrob{G}$ and $S_G$ into a commutative diagram,
determining the kernel and cokernel of $\TrFrob{G}$ when $G$ is an \'etale commutative group scheme over $\Fq$.

%
We begin with a simple, general result relating duals, invariants and coinvariants.

@@

Now let $\beta' \in Z^1(\Weil{},\bG^*)$ be another representative for $[\beta]$
so $\beta-\beta' = d_{\Weil{}} \delta$ for some $\delta \in C^0(\Weil{},\bG^*)$ defining $d \in \Hom(G(\bFq),\EEx)$.
Let $\cs{L}'$ be the character sheaf on $G$ defined by $\beta'$, as above.
For each $x\in G(\bFq)$, define $\brho_x :\cs{L}_x\to \cs{L}'_x$ by $\brho_x(1) = d(x)$.
The collection of isomorphisms $\{ \brho_x \tq x\in G(\bFq)\}$ satisfies condition~\ref{CS.4},
so it defines a morphism $\rho : \cs{L}\to \cs{L}'$, which is clearly an isomorphism. 
%
We have now defined a section of $\TrFrob{G}$.

@@

\begin{remark} \label{rem:H2}
Recall the K\"unneth formula in group cohomology \cite{brown:CohomologyGrps}*{Prop. I.0.8}:
if $A$ and $A'$ are groups and $M$ and $M'$ are abelian groups with $M$ $\ZZ$-free, then
\[
\Hh^n(A \times A', M \otimes M') \cong \bigoplus_{i+j=n} \Hh^i(A, M) \otimes \Hh^j(A', M') \ \ \oplus
\bigoplus_{i+j=n+1} \Tor_1^{\ZZ}\left(\Hh^i(A, M), \Hh^j(A', M')\right).
\]
%
Now suppose $\bG = \ZZ^r \times \prod_{i = 1}^m \ZZ/N_i\ZZ$ is an arbitrary finitely generated abelian group,
with $N_i \mid N_{i+1}$.  Then the K\"unneth formula implies that
\begin{equation} \label{eq:H2comp}
\Hh^2(\bG, \EEx) \cong \left(\EEx\right)^{r(r-1)/2} \times \prod_{i=1}^m (\ZZ/N_i\ZZ)^{m+r-i}.
\end{equation}
We see that $\Hh^2(\bG, \EEx)$ is trivial if and only if $\bG$ is cyclic.  Of course, $\Hh^0(\Weil{}, \Hh^2(\bG, \EEx))$
may or may not be trivial, even when $\Hh^2(\bG, \EEx)$ is non-trivial.
\end{remark}

@@

%While most of this paper focuses on the question of determining the isomorphism classes of objects, 
A complete understanding of the morphisms in $\CS(G)$ also requires a description of the
automorphisms of an arbitrary character sheaf $\cs{L}$.

@@

\begin{proof}
  By Lemma~\ref{lem:G0alg-grp} and Proposition~\ref{prop:connected}, every
  character sheaf on $G^0$ is isomorphic to $(\pi_! \EE)_\psi$ for some discrete isogeny $\pi : B \to G^0$ and character $\psi : \ker \pi \to \EEx$.
  So to prove the proposition it suffices to show that $(\pi_! \EE)_\psi$ extends to a character sheaf on $G$.
%
 By Lemma~\ref{lemma:ext}, there is an extension of the
 discrete isogeny $\pi : B \to G^0$ to a discrete isogeny $f : H \to G$
 such that $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism.
 Then $(f_! \EE)_\psi$ is a character sheaf on $G$ and
 $(f_! \EE)_\psi\vert_{G^0} \iso (\pi_! \EE)_\psi$.
\end{proof}

@@

As above, set $\bG^x \ceq \pi_0^{-1}(x)$ for $x\in \pi_0(\bG)$.
Let $g, g'$ be geometric points in the same
geometric connected component $\bG^x$.
Set $a = g^{-1}g'$ and note that $a$ is a geometric point in $\bG^0$.
Let $\mu_{g,a} : \gcs{L}_{ga} \to \gcs{L}_g \otimes \gcs{L}_a$
be the isomorphism of vector spaces obtained by restriction of
$\mu : m^*\gcs{L} \to \gcs{L} \boxtimes \gcs{L}$ to the
geometric point $(g,a)$ on $\bG^x \times \bG^0$.
Since $\cs{L}\vert_{G^0} = (\EE)_{G^0}$,
the stalk of $\gcs{L}$ at $a$ is $\EE$.
In this way the pair of geometric points $g, g' \in \bG^x$
determines an isomorphism $\varphi_{g,g'} \ceq \mu_{g,a}^{-1}$
from $\gcs{L}_{g}$ to $\gcs{L}_{g'}$.
%
The isomorphisms $\varphi_{g,g'}: \gcs{L}_{g} \to \gcs{L}_{g'}$ are canonical
in the following sense: if $g,g'\in \bG^x$ and $h,h'\in \bG^y$
then it follows from \ref{CS.2} and \ref{CS.3}
that
 \begin{equation}\label{eq:CS}
  \begin{tikzcd}[column sep=40]
   \gcs{L}_{gh} \arrow{r}{\varphi_{gh,g'h'}} \arrow[swap]{d}{\mu_{g,h}}
  & \gcs{L}_{g'h'} \arrow{d}{\mu_{g',h'}}
  &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}
  &  \gcs{L}_{\Frob{}(g)} \arrow{r}{\varphi_{\Frob{}(g),\Frob{}(g')}} \arrow[swap]{d}{\phi_{g}} & \gcs{L}_{\Frob{}(g')} \arrow{d}{\phi_{g'}} \\
  \gcs{L}_{g} \otimes \gcs{L}_{h} \arrow{r}{\varphi_{g,g'}\otimes \varphi_{h,h'}}
  & \gcs{L}_{g'} \otimes \gcs{L}_{h'}
  & {}
  & \gcs{L}_{g} \arrow{r}{\varphi_{g,g'}} & \gcs{L}_{g'}
  \end{tikzcd}
 \end{equation}
both commute.

@@

\begin{proof}
  Let
  \begin{equation}\label{eq:pre-snake}
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar
    &  G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0
  \end{tikzcd}
  \end{equation}
  be the commutative diagram of abelian groups obtained by applying
  Lemma~\ref{lem:pullback} to \eqref{eq:pi0}.
 %
  The sequence of abelian groups
\[
  \begin{tikzcd}
    1 \rar & G^0(\Fq) \rar & G(\Fq) \rar & \pi_0(G)(\Fq) \rar & 0,
  \end{tikzcd}
\]
  is exact since $\Hh^1(\Fq,G^0) =0$ by Lemma~\ref{lem:G0alg-grp} and Lang's theorem on connected algebraic groups over finite fields \cite{lang:56a}.
  Since $\EEx$ is divisible, $\Hom(\ - \ ,\EEx)$ is exact and thus the dual sequence of
  character groups in \eqref{eq:pre-snake} is exact.
%
  The upper row in \eqref{eq:pre-snake} is exact by Proposition~\ref{prop:middleexact}.
  Now Lemma~\ref{lem:G0alg-grp} and Proposition~\ref{prop:connected}
  imply that $\ker \TrFrob{G^0} =0$ and $\coker \TrFrob{G^0}=0$,
  while Proposition~\ref{prop:etale-iso} gives $\ker \TrFrob{\pi_0(G)} \iso \Hh^0(\Weil{},\Hh^2(\pi_0(\bG),\EEx))$
  and $\coker \TrFrob{\pi_0(G)}=0$.
It now follows from the snake lemma
 \begin{equation}\label{eq:snake}
  \begin{tikzcd}[row sep=20, column sep=15]
    0 \arrow[dotted]{r} & \ker \TrFrob{\pi_0(G)} \arrow[dotted]{r} \dar & \arrow{d} \ker \TrFrob{G} \arrow[dotted]{r}
    & \ker \TrFrob{G^0} =0 \arrow[dotted, out=-10, in=170]{dddll} \dar & \\
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
   &  \coker \TrFrob{\pi_0(G)} =0 \arrow[dotted]{r} & \coker \TrFrob{G} \arrow[dotted]{r} & \coker \TrFrob{G^0} =0 & 
  \end{tikzcd}
  \end{equation}
that $\coker \TrFrob{G} =0$
and $\ker \TrFrob{\pi_0(G)} \to \ker \TrFrob{G}$ is an isomorphism.
This gives the promised short exact sequence
\[
  \begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0. 
  \end{tikzcd}
\]
\end{proof}

@@

\begin{proposition} \label{prop:bounded}
Let $G$ be a smooth commutative group scheme over $\Fq$. 
Then $\cs{L}\in \CS(G)$ descends to $G$ if and only if $\trFrob{\cs{L}} : G(\Fq) \to \EEx$ has bounded image.
\end{proposition}
\begin{proof} 
By Lemma~\ref{lem:G0alg-grp}, the identity component $G^0$ is a connected algebraic group over $\Fq$. 
It follows from Proposition~\ref{prop:connected} that the restriction of $\cs{L}$ to $G^0$ descends to $G$. 
Also, since $G^0(\Fq)$ is finite, the image of $\trFrob{\cs{L}} : G(\Fq) \to \EEx$ is a finite subgroup and therefore has bounded image.  
%
If $\chi \in G(\Fq)^*$ then there is some finite-image character $\chi_0$
with the same restriction to $G^0(\Fq)$ since $G^0(\Fq)$ is lies inside the torsion part of
the finitely generated abelian group $G(\Fq)$.  Therefore $\chi$ is bounded 
if and only if $\chi \cdot \chi_0^{-1}$ is bounded.  But $\chi \cdot \chi_0^{-1}$ descends
to a character of $\pi_0(G)$.
Thus, it is enough to prove Corollary~\ref{prop:bounded}
for \'etale group schemes $G$, which is done in Proposition~\ref{prop:bounded-etale}.
\end{proof}

@@

Let $K$, $R$ and $R_n$ be as above.
For each $n \in \NN$, the Greenberg functor maps schemes over $R_n$ to schemes over $\Fq$.
See \cite{bertapelle-gonzales:Greenberg} for the definition and fundamental properties of the
Greenberg functor as we use it; other useful references include
\cite{greenberg:61}, \cite{greenberg:63a},
\cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1},
\cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} and
\cite{nicaise-sebag:motivicSerre}*{\S 2.2}. %citelist
%
For any non-negative integer $n$ we will write
\[
\Gr^R_n : \Sch{R} \to \Sch{\Fq}
\]
for the functor produced by composing pullback along $\Spec{R_n} \to \Spec{R}$ with the Greenberg functor. 
This functor respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and
geometric components.  Moreover, there is a canonical isomorphism
\[
\Gr^R_n(X)(\Fq) \iso X(R_n)
\]
for any scheme $X$ over $R$.

@@

\subsection{Quasicharacter sheaves}
%{Character sheaves on the Greenberg transform of a N\'eron model} 
\label{ssec:CS_on_GN}

%Let $X \in \mathcal{N}$ be as above.
Set $S = \Spec{R}$ and $S_n = \Spec{R_n}$;
note that $S_0 = \Spec{\Fq}$ is the special fibre of $S$.
Let $X$ be a smooth commutative group scheme over $S$.
%We now define a category which will geometrize quasicharacters of $X(K)$.
For every non-negative integer $n$, the Greenberg transform $\Gr^R_n(X)$ is a {\it smooth} commutative group scheme over $S_0$.
%with component group scheme $\pi_0(X)_{\Fq} \ceq \pi_0(X) \times_S S_0$ and $\Fq$-rational points $X(R_n).$
%
The Greenberg transform $\Gr_R(X)$ of $X$ is a commutative group scheme over $\Fq$
%with component group scheme $\pi_0(X)_{\Fq}$
with $\Fq$-rational points $X(R)$.
%
The morphism of $\Fq$-schemes $\varrho^R_n(X) : \Gr_R(X) \to \Gr^R_n(X)$ induces a functor
\[
\varrho^R_n(X)^* : \CS(\Gr^R_n(X)) \to \CS(\Gr_R(X)),
\]
as in Lemma~\ref{lem:pullback}.
%This functor is faithful, so we may view $\CS(\Gr^R_n(X))$ as a subcategory of $\CS(\Gr_R(X))$.

>>>>

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \labitem{(1)}{c1} $\TrFrob{G} : \CSiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \labitem{(2)}{c2} every character sheaf on $G$ is isomorphic to one defined by a discrete isogeny;
 \labitem{(3)}{c3} $\Aut(\cs{L}) = 1$, for all character sheaves $\cs{L}$ on $G$.
 \end{enumerate}
 \end{proposition}
\begin{proof}
By Lemma~\ref{lem:bounded_connected}, we know that every character sheaf $\cs{L}$ on $\bG$ descends to $G$;
let $\cs{E}$ be an object of $\bCS(G)$ for which $B_G(\cs{E}) \iso \cs{L}$.
 Since the functor $B_G : \bCS(G) \to \CS(G)$ is full and faithful, $\Aut(\cs{L}) = \Aut(\cs{E})$.
From here, Deligne's function-sheaf dictionary for connected commutative algebraic groups over finite fields,
as in \cite{deligne:SGA4.5}*{Sommes trig.} or \cite{laumon:87a}*{1.1.3}, gives us all we need for points \ref{c1} and \ref{c2}.
\iffalse
As in the proof of Proposition~\ref{prop:finite}, use \'etale descent to see that pullback by the Lang isogeny $\Lang : G\to G$
defines an equivalence of categories between local systems on $G$ and $G(\Fq)$-equivariant local systems on $G$. 
Under this equivalence, local systems $\cs{E}$ on $G$ arising from objects in $\bCS(G)$ are matched with $G(\Fq)$-equivariant
constant local systems of rank-one on $G$, and therefore with one-dimensional representations of $G(\Fq)$. 
In the same way, pullback along the isogeny $\Lang\times\Lang : G\times G\to G\times G$ matches the extra structure
$\mu_\cs{E} : m^*\cs{E} \to \cs{E}\boxtimes\cs{E}$ with an isomorphism $m^*V \to V\boxtimes V$ of one-dimensional
representations of $G(\Fq)\times G(\Fq)$, which is exactly an isomorphism $V \to V\otimes V$ of one-dimensional representations,
which is exactly the choice of an isomorphism $V\iso \EE$.
We see that $\bCS(G)$ is equivalent to the category of characters of $G(\Fq)$.
Let $\bg$ be a geometric point above $g \in G(\Fq)$.  If $\cs{E}$ matches $\psi : G(\Fq)\to \EEx$ under this equivalence,
a simple calculation on stalks reveals that the action of Frobenius on $\cs{E}_{\bg}$ is multiplication by $\psi(g)^{-1}$.
In other words, for every $\cs{E}$ in $\bCS(G)$, the trace of $\Lang^*\cs{E}$ is $\trFrob{\cs{E}}^{-1}$ as a
representation of $G(\Fq)$, proving parts \ref{c1} and \ref{c2}.
\fi
For part \ref{c3}, suppose $\Lang^*\cs{E} = V$ with isomorphism $V \to V\otimes V$.
Observe that the equivalence above establishes a bijection between $\Aut(\cs{E})$ and the group of automorphisms of $\rho : V\to V$ for which 
\[
\begin{tikzcd}
\arrow{d}{} V \arrow{r}{\rho} & V\arrow{d}{}\\
V\otimes V \arrow{r}{\rho\otimes \rho} & V\otimes V
\end{tikzcd}
\]
commutes. 
Since the only such isomorphism $\rho$ is $\id_V$, it follows that $\Aut(\cs{E}) = 1$, completing the proof.
\end{proof}

@@

@@

In this section we fit the group homomorphisms $\TrFrob{G}$ and $S_G$ into a commutative diagram,
determining the kernel and cokernel of $\TrFrob{G}$ when $G$ is an \'etale commutative group scheme over $\Fq$.
We begin with a simple, general result relating duals, invariants and coinvariants.

@@

Now let $\beta' \in Z^1(\Weil{},\bG^*)$ be another representative for $[\beta]$
so $\beta-\beta' = d_{\Weil{}} \delta$ for some $\delta \in C^0(\Weil{},\bG^*)$ defining $d \in \Hom(G(\bFq),\EEx)$.
Let $\cs{L}'$ be the character sheaf on $G$ defined by $\beta'$, as above.
For each $x\in G(\bFq)$, define $\brho_x :\cs{L}_x\to \cs{L}'_x$ by $\brho_x(1) = d(x)$.
The collection of isomorphisms $\{ \brho_x \tq x\in G(\bFq)\}$ satisfies condition~\ref{CS.4},
so it defines a morphism $\rho : \cs{L}\to \cs{L}'$, which is clearly an isomorphism. 
We have now defined a section of $\TrFrob{G}$.

@@

\begin{remark} \label{rem:H2}
Recall the K\"unneth formula in group cohomology \cite{brown:CohomologyGrps}*{Prop. I.0.8}:
if $A$ and $A'$ are groups and $M$ and $M'$ are abelian groups with $M$ $\ZZ$-free, then
\[
\Hh^n(A \times A', M \otimes M') \cong \bigoplus_{i+j=n} \Hh^i(A, M) \otimes \Hh^j(A', M') \ \ \oplus
\bigoplus_{i+j=n+1} \Tor_1^{\ZZ}\left(\Hh^i(A, M), \Hh^j(A', M')\right).
\]
Now suppose $\bG = \ZZ^r \times \prod_{i = 1}^m \ZZ/N_i\ZZ$ is an arbitrary finitely generated abelian group,
with $N_i \mid N_{i+1}$.  Then the K\"unneth formula implies that
\begin{equation} \label{eq:H2comp}
\Hh^2(\bG, \EEx) \cong \left(\EEx\right)^{r(r-1)/2} \times \prod_{i=1}^m (\ZZ/N_i\ZZ)^{m+r-i}.
\end{equation}
We see that $\Hh^2(\bG, \EEx)$ is trivial if and only if $\bG$ is cyclic.  Of course, $\Hh^0(\Weil{}, \Hh^2(\bG, \EEx))$
may or may not be trivial, even when $\Hh^2(\bG, \EEx)$ is non-trivial.
\end{remark}

@@

A complete understanding of the morphisms in $\CS(G)$ also requires a description of the
automorphisms of an arbitrary character sheaf $\cs{L}$.

@@

\begin{proof}
  By Lemma~\ref{lem:G0alg-grp} and Proposition~\ref{prop:connected}, every
  character sheaf on $G^0$ is isomorphic to $(\pi_! \EE)_\psi$ for some discrete isogeny $\pi : B \to G^0$ and character $\psi : \ker \pi \to \EEx$.
  So to prove the proposition it suffices to show that $(\pi_! \EE)_\psi$ extends to a character sheaf on $G$.
 By Lemma~\ref{lemma:ext}, there is an extension of the
 discrete isogeny $\pi : B \to G^0$ to a discrete isogeny $f : H \to G$
 such that $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism.
 Then $(f_! \EE)_\psi$ is a character sheaf on $G$ and
 $(f_! \EE)_\psi\vert_{G^0} \iso (\pi_! \EE)_\psi$.
\end{proof}

@@

As above, set $\bG^x \ceq \pi_0^{-1}(x)$ for $x\in \pi_0(\bG)$.
Let $g, g'$ be geometric points in the same
geometric connected component $\bG^x$.
Set $a = g^{-1}g'$ and note that $a$ is a geometric point in $\bG^0$.
Let $\mu_{g,a} : \gcs{L}_{ga} \to \gcs{L}_g \otimes \gcs{L}_a$
be the isomorphism of vector spaces obtained by restriction of
$\mu : m^*\gcs{L} \to \gcs{L} \boxtimes \gcs{L}$ to the
geometric point $(g,a)$ on $\bG^x \times \bG^0$.
Since $\cs{L}\vert_{G^0} = (\EE)_{G^0}$,
the stalk of $\gcs{L}$ at $a$ is $\EE$.
In this way the pair of geometric points $g, g' \in \bG^x$
determines an isomorphism $\varphi_{g,g'} \ceq \mu_{g,a}^{-1}$
from $\gcs{L}_{g}$ to $\gcs{L}_{g'}$.
The isomorphisms $\varphi_{g,g'}: \gcs{L}_{g} \to \gcs{L}_{g'}$ are canonical
in the following sense: if $g,g'\in \bG^x$ and $h,h'\in \bG^y$
then it follows from \ref{CS.2} and \ref{CS.3}
that
 \begin{equation}\label{eq:CS}
  \begin{tikzcd}[column sep=40]
   \gcs{L}_{gh} \arrow{r}{\varphi_{gh,g'h'}} \arrow[swap]{d}{\mu_{g,h}}
  & \gcs{L}_{g'h'} \arrow{d}{\mu_{g',h'}}
  &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}
  &  \gcs{L}_{\Frob{}(g)} \arrow{r}{\varphi_{\Frob{}(g),\Frob{}(g')}} \arrow[swap]{d}{\phi_{g}} & \gcs{L}_{\Frob{}(g')} \arrow{d}{\phi_{g'}} \\
  \gcs{L}_{g} \otimes \gcs{L}_{h} \arrow{r}{\varphi_{g,g'}\otimes \varphi_{h,h'}}
  & \gcs{L}_{g'} \otimes \gcs{L}_{h'}
  & {}
  & \gcs{L}_{g} \arrow{r}{\varphi_{g,g'}} & \gcs{L}_{g'}
  \end{tikzcd}
 \end{equation}
both commute.

@@

\begin{proof}
  Let
  \begin{equation}\label{eq:pre-snake}
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar
    &  G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0
  \end{tikzcd}
  \end{equation}
  be the commutative diagram of abelian groups obtained by applying
  Lemma~\ref{lem:pullback} to \eqref{eq:pi0}.
  The sequence of abelian groups
\[
  \begin{tikzcd}
    1 \rar & G^0(\Fq) \rar & G(\Fq) \rar & \pi_0(G)(\Fq) \rar & 0,
  \end{tikzcd}
\]
  is exact since $\Hh^1(\Fq,G^0) =0$ by Lemma~\ref{lem:G0alg-grp} and Lang's theorem on connected algebraic groups over finite fields \cite{lang:56a}.
  Since $\EEx$ is divisible, $\Hom(\ - \ ,\EEx)$ is exact and thus the dual sequence of
  character groups in \eqref{eq:pre-snake} is exact.
  The upper row in \eqref{eq:pre-snake} is exact by Proposition~\ref{prop:middleexact}.
  Now Lemma~\ref{lem:G0alg-grp} and Proposition~\ref{prop:connected}
  imply that $\ker \TrFrob{G^0} =0$ and $\coker \TrFrob{G^0}=0$,
  while Proposition~\ref{prop:etale-iso} gives $\ker \TrFrob{\pi_0(G)} \iso \Hh^0(\Weil{},\Hh^2(\pi_0(\bG),\EEx))$
  and $\coker \TrFrob{\pi_0(G)}=0$.
It now follows from the snake lemma
 \begin{equation}\label{eq:snake}
  \begin{tikzcd}[row sep=20, column sep=15]
    0 \arrow[dotted]{r} & \ker \TrFrob{\pi_0(G)} \arrow[dotted]{r} \dar & \arrow{d} \ker \TrFrob{G} \arrow[dotted]{r}
    & \ker \TrFrob{G^0} =0 \arrow[dotted, out=-10, in=170]{dddll} \dar & \\
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
   &  \coker \TrFrob{\pi_0(G)} =0 \arrow[dotted]{r} & \coker \TrFrob{G} \arrow[dotted]{r} & \coker \TrFrob{G^0} =0 & 
  \end{tikzcd}
  \end{equation}
that $\coker \TrFrob{G} =0$
and $\ker \TrFrob{\pi_0(G)} \to \ker \TrFrob{G}$ is an isomorphism.
This gives the promised short exact sequence
\[
  \begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0. 
  \end{tikzcd}
\]
\end{proof}

@@

\begin{proposition} \label{prop:bounded}
Let $G$ be a smooth commutative group scheme over $\Fq$. 
Then $\cs{L}\in \CS(G)$ descends to $G$ if and only if $\trFrob{\cs{L}} : G(\Fq) \to \EEx$ has bounded image.
\end{proposition}
\begin{proof} 
By Lemma~\ref{lem:G0alg-grp}, the identity component $G^0$ is a connected algebraic group over $\Fq$. 
It follows from Proposition~\ref{prop:connected} that the restriction of $\cs{L}$ to $G^0$ descends to $G$. 
Also, since $G^0(\Fq)$ is finite, the image of $\trFrob{\cs{L}} : G(\Fq) \to \EEx$ is a finite subgroup and therefore has bounded image.  
If $\chi \in G(\Fq)^*$ then there is some finite-image character $\chi_0$
with the same restriction to $G^0(\Fq)$ since $G^0(\Fq)$ is lies inside the torsion part of
the finitely generated abelian group $G(\Fq)$.  Therefore $\chi$ is bounded 
if and only if $\chi \cdot \chi_0^{-1}$ is bounded.  But $\chi \cdot \chi_0^{-1}$ descends
to a character of $\pi_0(G)$.
Thus, it is enough to prove Corollary~\ref{prop:bounded}
for \'etale group schemes $G$, which is done in Proposition~\ref{prop:bounded-etale}.
\end{proof}

@@

Let $K$, $R$ and $R_n$ be as above.
For each $n \in \NN$, the Greenberg functor maps schemes over $R_n$ to schemes over $\Fq$.
See \cite{bertapelle-gonzales:Greenberg} for the definition and fundamental properties of the
Greenberg functor as we use it; other useful references include
\cite{greenberg:61}, \cite{greenberg:63a},
\cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1},
\cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} and
\cite{nicaise-sebag:motivicSerre}*{\S 2.2}.
For any non-negative integer $n$ we will write
\[
\Gr^R_n : \Sch{R} \to \Sch{\Fq}
\]
for the functor produced by composing pullback along $\Spec{R_n} \to \Spec{R}$ with the Greenberg functor. 
This functor respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and
geometric components.  Moreover, there is a canonical isomorphism
\[
\Gr^R_n(X)(\Fq) \iso X(R_n)
\]
for any scheme $X$ over $R$.

@@

\subsection{Quasicharacter sheaves}
\label{ssec:CS_on_GN}

Set $S = \Spec{R}$ and $S_n = \Spec{R_n}$;
note that $S_0 = \Spec{\Fq}$ is the special fibre of $S$.
Let $X$ be a smooth commutative group scheme over $S$.
For every non-negative integer $n$, the Greenberg transform $\Gr^R_n(X)$ is a {\it smooth} commutative group scheme over $S_0$.
The Greenberg transform $\Gr_R(X)$ of $X$ is a commutative group scheme over $\Fq$
with $\Fq$-rational points $X(R)$.
The morphism of $\Fq$-schemes $\varrho^R_n(X) : \Gr_R(X) \to \Gr^R_n(X)$ induces a functor
\[
\varrho^R_n(X)^* : \CS(\Gr^R_n(X)) \to \CS(\Gr_R(X)),
\]
as in Lemma~\ref{lem:pullback}.
